Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-02-10 cluster_paper/ms.tex
update paper

%\newcommand{\nautorot}{1{,}144\ } % % For nequal<=0, nclose<=100, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorot}{987\ } % % For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt

@@

\newcommand{\ncalibration}{9{,}619{,}784\ } % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\nnbhdcaliblc}{10{,}314\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\nnbhdcalibstar}{2{,}212\ } % make_compstar_NGC_2516_sourcelist.py

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these ``halos'' could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a 300\,pc halo reported
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
  tidal radius of the cluster is $\sim$10\,pc.  Using photometry
  from Gaia, rotation periods from TESS, and lithium measurements from
  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
  coeval with the core.  Two in three kinematically selected halo
  members 25-250\,pc from the cluster center show rotation periods
  consistent with a gyrochronological age of 120\,Myr.  Stars in the
  inner 25\,pc show higher period detection rates than the
  halo (73\% vs.\ 66\%), but both vastly exceed our field star
  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
  are also anti-correlated at fixed stellar mass, as is the case with
  comparable clusters.  This work expands the set of confirmed
  NGC\,2516 members by a factor of $\approx$2, and quantifies the
  degree to which Gaia-based analyses can identify dispersed stellar
  populations against a background of field star contaminants.  We
  highlight implications for spectroscopic survey targeting, open
  cluster dispersal, and planet searches around young stars.
\end{abstract}
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023

@@

@@

Figure~\ref{fig:rot} shows the resulting rotation periods.  The points
on this plot are the \nautorot light curves for which the peak Lomb Scargle
periodogram period was below 15 days, the LS power exceeded 0.08, and
for which no equal-brightness companions were within the aperture.  These
selection criteria are entirely heuristic.  
But splitting the sample into the ``core'' and the ``halo'', they
suggest that the halo is real.

@@

\paragraph{Comparison sample}
To provide a basis for comparison, we also opted to search the
``calibration'' light curves ($Rp<13$) that were created as a part of
the CDIPS project. 
Over the southern sky (Sectors 1-13 of TESS), this corresponded to a sample of \ncalibration stars.
Cross-matching these against \nnbhd randomly drawn stars in the
neighborhood of \cn yielded \nnbhdcalibstar unique stars, with
\nnbhdcaliblc sector-specific TESS light curves.

@@

%/Users/luke/Dropbox/proj/earhart/results/lithium/kinematic_X_galah_dr3.csv
% https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
We downloaded\footnote{Via \url{datacentral.org.au/services/download},
using the \texttt{sobject_id} identifiers.}
the GALAH DR3 spectra for all 106 entries.
We then measured the lithium equivalent widths using PROCESS Y.
We then concatenated the GALAH DR3 and Gaia-ESO results together.

>>>>

%\newcommand{\nautorot}{1{,}144\ } % % For nequal<=0, nclose<=100, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt

@@

\newcommand{\ncalibration}{9{,}619{,}784\ } % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\nnbhdcaliblc}{9{,}790\ } % get_auto_rotation_periods.py
\newcommand{\nnbhdcalibstar}{1{,}987\ } % get_auto_rotation_periods.py

\newcommand{\nautorotdenominator}{1{,}641\ } % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987\ } % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\ncompstardenominator}{820\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\ncompstarnumerator}{365\ } % plot_autorotation.py
\newcommand{\ncompfrac}{44.5\%} % plot_autorotation.py
\newcommand{\nautovscompstardenominator}{705\ } % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374\ } % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these halos could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a halo with a length of 300\,pc reported around
  the $\approx$120\,Myr open cluster NGC\,2516.  The classical tidal
  radius of the cluster is $\sim$10\,pc.  Using photometry from Gaia,
  rotation periods from TESS, and lithium measurements from Gaia-ESO
  and GALAH, we find that the halo of NGC\,2516 is real and coeval
  with the core.  Two in three kinematically selected halo members
  25-250\,pc from the cluster center show rotation periods consistent
  with a gyrochronological age of 120\,Myr.  A comparison sample of
  field stars shows no such trend.  The rotation period and lithium
  abundance for the cluster stars are also anti-correlated at fixed
  stellar mass, as is the case with comparable stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.  We highlight implications
  for spectroscopic survey targeting, open cluster dispersal, and
  planet searches around young stars.
\end{abstract}

@@

\subsubsection{Cluster Star Sample}

@@

Figure~\ref{fig:rot} shows the resulting rotation periods.  The points
on this plot are the \nautorotnumerator light curves for which the peak Lomb Scargle
periodogram period was below 15 days, the LS power exceeded 0.08, and
for which no equal-brightness or greater companions were within the aperture. 
We also required that at most one companion with brightness exceeding 
one tenth of the target star could reside in the aperture.
These
selection criteria are entirely heuristic, and yielded
\nautorotdenominator stars for which rotation periods might have been
detected.  
By exploring the different possible selection criteria, we found that
the ones we chose provided a
reasonable balance between excluding stars on the slow sequence, {\it
vs.} reporting spurious rotation periods in the ``void'' below the
slow sequence, where the dominant signal was in fact from a
neighboring star.
Splitting the resulting sample into the ``core'' and the ``halo'', they
suggest that the halo is real.

@@

\subsubsection{Comparison Star Sample}

\begin{figure}[t]
	\begin{center}
		\leavevmode
    \includegraphics[width=0.49\textwidth]{f9.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Field stars show a different rotation period
  distribution than the kinematically selected members.}
		\label{fig:compstar}
	}
\end{figure}

To provide a basis for comparison, we also opted to search the
``calibration'' light curves ($Rp<13$) that were created as a part of
the CDIPS project.  Over the southern sky (Sectors 1-13 of TESS), this
corresponded to a sample of \ncalibration stars.  Cross-matching these
against the \nnbhd randomly drawn stars in the neighborhood of \cn yielded
\nnbhdcalibstar unique stars, with a cumulative total of \nnbhdcaliblc TESS
sectors observed.
The magnitude cut of $Rp<13$ at the distances of the neighborhood sample
corresponds to an extinction-corrected color cutoff of $(Bp-Rp)_0
\approx 0.80$, or spectral types of $\approx$G1V.
This reaches sufficiently far down the main sequence to enable a
comparison to the cluster star sample.

We performed the same light curve stitching and period-search
procedure on the field comparison stars.
Imposing the same requirements for crowding resulted in
\ncompstardenominator stars for which rotation periods could have been
detected.
Imposing the same Lomb Scargle power cutoff, and period upper limit,
yielded \ncompstarnumerator period detections (\ncompfrac).
Within the same brightness cutoff, 
\nautovscompstarnumerator of \nautovscompstardenominator cluster stars
yielded period detections (\nautofrac).
Though the detection fractions are frankly not very different
(likely because of the brightness cutoff), the period {\it
vs} color distributions are quite different
(Figure~\ref{fig:compstar}).

@@

%/Users/luke/Dropbox/proj/earhart/results/lithium/kinematic_X_galah_dr3.csv
% https://docs.datacentral.org.au/galah/dr3/catalogue-data-access/
We downloaded\footnote{Via \url{datacentral.org.au/services/download},
using the \texttt{sobject\_id} identifiers.}
the GALAH DR3 spectra for all 106 entries.
We then measured the lithium equivalent widths using PROCESS Y.
We then concatenated the GALAH DR3 and Gaia-ESO results together.
